
\newglossaryentry{cmd:target}{
name={target},
text={Target Block},
description={Name of the destination node, equivalent to target of defdst edge in schedule.},
type=cmd}

\newglossaryentry{cmd:pattern}{type=cmd,
name={pattern},
text={Target Pattern},
description={See \gls{cmd:target}},
}

\newglossaryentry{cmd:dest}{
name={dest},
description={Name of the destination node, equivalent to target of defdst edge in schedule.},
type=cmd}

\newglossaryentry{cmd:destpattern}{type=cmd,
name={destpattern},
text={Destination Pattern},
description={Name of the destination pattern. Equivalent to \gls{cmd:dest} to the patterns entry node. },
}

\newglossaryentry{cmd:destbeamproc}{
name={destbeamproc},
text={Dest. beamprocess},
description={Name of the destination beamprocess. Equivalent to \gls{cmd:dest} to the beamprocess' entry node.},
type=cmd}

\newglossaryentry{cmd:permanent}{
name={permanent},
description={If true, the flow command will permanently change the default destination of the target block, else the change lasts only during the execution of the command.},
type=cmd}

\newglossaryentry{cmd:qil}{
name={qil},
description={Flush queue selection, flushes high priority queue if true},
type=cmd}

\newglossaryentry{cmd:qhi}{
name={qhi},
description={Flush queue selection, flushes mid priority queue if true},
type=cmd}

\newglossaryentry{cmd:qlo}{
name={qlo},
description={Flush queue selection, flushes low priority queue if true},
type=cmd}


\newglossaryentry{cmd:tvalid}{
name={tvalid},
text={Valid Time},
description={Time after which this command will be valid (executable), can be absolute or relative depending on \gls{cmd:vabs} flag},
type=cmd}

\newglossaryentry{cmd:vabs}{
name={vabs},
description={Chooses whether tvalid is interpreted as absolute value or an offset to current time sum.},
type=cmd}

\newglossaryentry{cmd:prio}{
name={prio},
text={Queue Priority},
description={Chooses the queue priority level this command will be written to at the target block (0 (lo), 1(mid), 2(hi))},
type=cmd}

\newglossaryentry{cmd:qty}{
name={qty},
description={Repetition quantity of a command. The sent command will be executed until qty reaches 0},
type=cmd}

\newglossaryentry{cmd:twait}{
name={twait},
description={Timespan a wait command is going to wait. Can absolute or relative to current time, depending on flag},
type=cmd}
