\newglossaryentry{sched:type}{
name={type},
description={Determines the type of this node (e.g. block, tmsg, etc.)},
type=sched}

\newglossaryentry{sched:type:tmsg}{
name={tmsg},
description={Timing Message. Causes a message to be broadcasted to all timing receivers},type=sched,
parent={sched:type}
}

\newglossaryentry{sched:type:block}{
name={block},
description={Fixed length block. Terminates a sequence of nodes and defines the sequence's length in time. See \gls{sched:tperiod}},
type=sched,
parent={sched:type}
}

\newglossaryentry{sched:type:blockalign}{
name={blockalign},
description={Alignment block. Extends its own length so the time sum will become a multiple of the time grid (currently \SI{10}{\micro\second}). See \gls{sched:tperiod}, \gls{sched:type:block}},
type=sched,
parent={sched:type}
}

\newglossaryentry{sched:type:flow}{
name={flow},
description={Causes one or more repetitions of a flow command to be written to the target block's queue, changing it's default successor (e.g. branch, loop, etc) },
type=sched,
parent={sched:type}
}

\newglossaryentry{sched:type:flush}{
name={flush},
description={Causes a flush command to be written to the target block's queue, emptying any selected queues. Can optionally also override the default successor},
type=sched,
parent={sched:type}
}

\newglossaryentry{sched:type:wait}{
name={wait},
description={Causes a wait command to be written to the target block's queue, temporarily stretching it's length in time},
type=sched,
parent={sched:type}
}

\newglossaryentry{sched:type:noop}{
name={nop},
description={Causes one or more No Operation command(s) to be written to the target block's queue. No direct effect, but can together with flow create a sequence of default and alternative successors},
type=sched,
parent={sched:type}
}



\newglossaryentry{sched:node_id}{
name={node\_id},
description={Tag of the node's name. Internal use only},
type=sched}

\newglossaryentry{sched:cpu}{
name={cpu},
description={Index of the CPU core this Node will reside in. To be replaced by auto-balancer algorithm},
type=sched}

\newglossaryentry{sched:thread}{
name={thread},
description={Index of the thread core this Node will be handled by. To be replaced by auto-balancer algorithm},
type=sched}

\newglossaryentry{sched:flags}{
name={flags},
description={Aggregation of certain node flags, internal use only},
type=sched}

\newglossaryentry{sched:patentry}{
name={patentry},
description={If true, this node is an entry point to the pattern it belongs to},
type=sched}

\newglossaryentry{sched:patexit}{
name={patexit},
description={If true, this node is an exit point from the pattern it belongs to},
type=sched}

\newglossaryentry{sched:pattern}{
name={pattern},
description={The name of thepattern this node belongs to},
type=sched}

\newglossaryentry{sched:bpentry}{
name={bpentry},
description={If true, this node is an entry point to the beamprocess it belongs to. Currently not supported},
type=sched}

\newglossaryentry{sched:bpexit}{
name={bpexit},
description={If true, this node is an exit point from the beamprocess it belongs to. Currently not supported},
type=sched}

\newglossaryentry{sched:beamproc}{
name={beamproc},
description={The name of the beamprocess this node belongs to. Currently not supported},
type=sched}

\newglossaryentry{sched:permanent}{
name={permanent},
description={If true, the corresponding flow command will permanently change the default destination of the target block, else the change lasts only during the execution of the command},
type=sched}

\newglossaryentry{sched:tperiod}{
name={tperiod},
description={The duration of a block. Amount of time in \SI{}{\nano\second} it adds to the current time sum when processed},
type=sched}

\newglossaryentry{sched:qil}{
name={qil},
description={Block attribute, generates a command queue of priority High if true},
type=sched}

\newglossaryentry{sched:qhi}{
name={qhi},
description={Block attribute, generates a command queue of priority Mid if true},
type=sched}

\newglossaryentry{sched:qlo}{
name={qlo},
description={Block attribute, generates a command queue of priority Low if true},
type=sched}

\newglossaryentry{sched:toffs}{
name={toffs},
description={The time offset \SI{}{\nano\second} relative to the block start at which this node is executed. Negative offsets can be used for debugging to force late events},
type=sched}

\newglossaryentry{sched:id}{
name={id},
description={Addresses the 64b ID field of a timing event. Sub-Id fields can be used instead, see \gls{id:id}},
type=sched}

\newglossaryentry{sched:tvalid}{
name={tvalid},
description={Time in \SI{}{\nano\second} at/after which this command will be valid (executable), can be absolute or relative depending on \gls{sched:vabs} flag},
type=sched}

\newglossaryentry{sched:vabs}{
name={vabs},
description={Chooses whether tvalid is interpreted as absolute value or an offset to current time sum. When using this as a loop initialiser head, \emph{always} use vabs=true and tvalid=0!},
type=sched}

\newglossaryentry{sched:prio}{
name={prio},
description={Chooses the command queue level this command will be written to at the target block. (0 (lo), 1(mid), 2(hi) )},
type=sched}

\newglossaryentry{sched:qty}{
name={qty},
description={Repetition quantity of a command(-node). The generated command will be executed until qty reaches 0},
type=sched}

\newglossaryentry{sched:twait}{
name={twait},
description={Timespan in \SI{}{\nano\second} a wait command is goid to wait. Can absolute or relative to current time, depending on flag},
type=sched}

\newglossaryentry{sched:par}{
name={par},
description={Transparent parameter field, part of a timing message. Is passed through to a listening ECA channel in timing receiver},
type=sched}

\newglossaryentry{sched:tef}{
name={tef},
description={Fractional (sub-nanosecond) time extension field, part of a timing message. Currently not interpreted by ECA)},
type=sched}

