\documentclass[12pt,a4paper]{report}
% Language: English
\pdfminorversion=7
\usepackage[pdftex]{graphicx}
\usepackage{changepage}
\usepackage{xcolor}
\usepackage{listings}

\lstdefinestyle{dotfiles}{
  escapeinside={(*@}{@*)}, % (*@\label{mylabel}@*)
  numbers=left,
  stepnumber=1,
  numberstyle=\tiny,
  numbersep=10pt,
  captionpos=b,
  belowcaptionskip=1\baselineskip,
  breaklines=true,
  keepspaces=true,
  columns=flexible,
  language=C,
  showstringspaces=false,
  basicstyle=\scriptsize\ttfamily,
  keywordstyle=\color{green!40!black},
  commentstyle=\itshape\color{purple!40!black},
  identifierstyle=\color{blue},
  stringstyle=\color{red},
  tabsize=2,
  morekeywords={digraph, graph, subgraph, edge, node, color, style, shape, fillcolor},
}

\newcommand{\ry}{\rotatebox{90}}
\begin{document}

\begin{titlepage}
\vspace{2cm}
\begin{center}
\Huge{Tests for the Datamaster}

\Large{Martin Skorsky}

\Large{Last change: 2021-03-17}
\end{center}
\vfill
\end{titlepage}

\tableofcontents

\chapter{Overview - What is tested}
All tests are on branch \texttt{dm-fallout-tests}. The tests run with \texttt{make} or \texttt{make all} in folder \texttt{modules/ftm/tests}.
To compile \texttt{libcarpedm} use \texttt{make prepare}. This runs \texttt{make clean} and \texttt{make} in folder \texttt{modules/ftm/ftmx86}.

Each test has a seperate target in the \texttt{Makefile}.
\begin{table}
\caption{Which test tests what}
%\begin{center}
\centering
\begin{tabular}[t]{|l|c|c|c|c|c|c|c|c|}
\hline
Test               & \ry{Tools} & \ry{libcarpedm} & \ry{firmware} & \ry{uses Python} & \ry{common} & \ry{make} & \ry{checks result} \\ \hline
addDownloadCompare &   T        &   T             &   T           &   x              &   x         &   -       &   -                \\ \hline
bpcStart           &   -        &   T             &   x           &   x              &   x         &   -       &   x                \\ \hline
dmCmd              &   T        &   x             &   x           &   x              &   -         &   -       &   x                \\ \hline
dmPerformance      &   x        &   T             &   x           &   -              &   -         &   x       &   -                \\ \hline
dmThreads          &   x        &   x             &   T           &   -              &   -         &   x       &   -                \\ \hline
fid7               &   -        &   x             &   T           &   x              &   x         &   -       &   x                \\ \hline
full\_test         &   -        &   T             &   x           &   x              &   -         &   -       &   x                \\ \hline
pps                &   -        &   T             &   x           &   x              &   -         &   -       &   -                \\ \hline
safe2remove        &   x        &   T             &   T           &   -              &   -         &   x       &   x                \\ \hline
singleEdgeTest     &   -        &   T             &   -           &   -              &   -         &   -       &   x                \\ \hline
schedules          &   x        &   x             &   T           &   -              &   -         &   x       &   -                \\ \hline
\end{tabular}
%\end{center}
\end{table}
\chapter{The Tests}
\section{addDownloadCompare}
\texttt{addDownloadCompare} test that a schedule is equivalent to the schedule which is downloaded form the datamaster firmaware.
The test steps are: clear the datamaster, add a schedule, start all pattern, download the schedule, compare both schedules with \texttt{scheduleCompare}. 
There should be no difference with \texttt{scheduleCompare}. Each test case uses a different schedule.
\section{bpcStart}
\texttt{bpcStart} tests the implementation of the beam process chain start flag in libcarpedm.
The test schedule sends two timing messages with \texttt{bpcstart=True} and \texttt{bpcstart=1}.
With \texttt{saft-ctl snoop} it is checked that the timing messages contain the correct setting.
In addition with \texttt{dm-sched} the dumped schedule is checked for the bpcstart flag.
\section{dmCmd}
\texttt{dmCmd} contains Python test scripts which contain Python unit tests for the tool dm-cmd. Each unit test calls dm-cmd
with commands and options and checks the result with the output on stdout and stderr. There are also negative tests with an invalid command
line. These tests are successful when the response is the correct error message and not a core dump.
\section{dmPerformance}
\texttt{dmPerformance} tests the performance improvements in libcarpedm.
The test starts a schedule on a clean data master, checks if some part of the schedule is removable, removes it and
then adds another schedule. This is done for a small schedule and a larger schedule. The test is ok if all commands
work. There is no check for this.
\section{dmThreads}
\texttt{dmThreads} tests the firmware with 4 threads. The same test for 6 threads fails due to a bug in the
firmware. For each thread a pattern with one block and one timing message is loaded.
\section{fid7}
\texttt{fid7} tests the fix for the format id 7 bug.
\include{full_test}
\section{messageCounter} This test is obsolete.
\texttt{messageCounter} is a basic test for the Python-Wrapper of libcarpedm.
\section{pps}
\texttt{pps} (pulse per second) is a basic test with a schedule which sends two timing messages every second.
\section{safe2remove}
\texttt{safe2remove} tests to remove a pattern from a running schedule. Test steps:
\begin{enumerate}
\item Clear data master
\item Add schedule
\item Start all patterns
\item Check removal of one pattern, should fail while pattern is running.
\item Abort pattern
\item Check removal of this pattern, should be valid, since pattern is not running.
\item Remove this pattern.
\item Check status of remaining schedule.
\end{enumerate}
These test steps are applied to a bunch of schedules.
\include{singleEdgeTest}
\section{schedules}
\texttt{schedules} collects schedules which are started. Tests that the schedules are compiled and loaded.
\chapter{Common Components}
\section{dm\_testbench.py}
\texttt{dm\_testbench.py} is a collection of Python functions for use in other test scripts.
\begin{enumerate}
\item startpattern(data\_master, pattern\_file)
    Connect to the given data master and load the pattern file (dot format).
    The data master is halted, cleared, and statistics is reset.
    Search for the first pattern in the data master with 'dm-sched' and start it.
\end{enumerate}
\end{document}
