\documentclass[12pt,a4paper]{report}
% Language: English
\usepackage{xurl}

\newcommand{\ry}{\rotatebox{90}}
\begin{document}

\begin{titlepage}
\vspace{2cm}
\begin{center}
\Huge{Documentation Data Master Test System}

\Large{Martin Skorsky}

\Large{Last change: 2023-10-10}
\end{center}
\vfill
\end{titlepage}

\tableofcontents

\chapter{Hardware}
\section{SuperMicro fel0069}
The data master for the test system is hosted on SuperMicro fel0069 with two PEXARIA5d \texttt{(fel0069.acc.gsi.de)}.
One of these is the data master. The other one is used to analyse the timing messages with snoop.
The SuperMicro is accessible with \texttt{ssh root@fel0069.acc.gsi.de} via ASL cluster or other hosts.
The management interface (ILO) is accessible via \url{https://fel0069i.acc.gsi.de}.
\begin{enumerate}
\item fel0069.acc.gsi.de
\begin{itemize}
\item IP: 140.181.169.146 (in Hex für PXE Boot: 8C.B5.A9.92)
\item MAC ILO: ac:1f:6b:71:fe:8f, IPv4 ILO: 140.181.169.147
\item MAC Lan1: ac:1f:6b:72:02:2c
\item MAC Lan2: ac:1f:6b:72:02:2d
\item Location: BG2.009, Rack BG2A.A9, Slot 28
\end{itemize}
\item pexaria248t (dev/wbm0), Data master
\begin{itemize}
\item IP: 192.168.131.184
\item MAC: 00:26:7b:00:08:0b
\item Name: pexaria248t
\item CID: 55 0113 0012 0
\item PEXARIA5d, Serie EE
\end{itemize}
\item pexaria305t (dev/wbm1), Timing receiver for snoop
\begin{itemize}
\item IP: 192.168.131.241
\item MAC: 00:26:7b:00:08:44
\item Name: pexaria305t
\item CID: 55 0113 0069 4
\item PEXARIA5d, Serie EE
\end{itemize}
\end{enumerate}
After power on, set the IP addresses of the two pexarias. This host has no BootP service.

The SuperMicro is configured for PXE boot and nfsinit with links (following
\url{https://www-acc.gsi.de/wiki/Timing/Intern/TimingSystemHowToHintsForFECS}).
\begin{enumerate}
\item On ASL cluster links in folder /common/tftp/csco/pxe/pxelinux.cfg for PXE boot:
8CB5A992 $\to$ fel0069,
fel0069 $\to$ scuxl.fallout
\item On ASL cluster links in folder /common/export/nfsinit/fel0069/ for nfsinit.
\item On ASL cluster create folder /common/fesadata/data/fel0069/.
\end{enumerate}
Connections: network cable to acc network. For management: network cable to acc network.

After \texttt{reboot} of fel0069:
\begin{enumerate}
\item on acopc042 (or some other host/user which needs access to fel0069) copy the public ssh key to fel0069.
This has to be done for every user / host which needs access to fel0069 with ssh.\raggedright
\linebreak ssh-copy-id -i .ssh/id\_rsa.pub root@fel0069.acc

\item test (check before automated testing starts)\raggedright
\linebreak ssh -t root@fel0069.acc.gsi.de "saft-ctl tr1 -xv snoop 0 0 0 3"\raggedright
\linebreak (snoop on tr1 for all events for 3 seconds)
\end{enumerate}

\section{White Rabbit Switch nwt0473m66}
Location: BG2.009, Rack BG2A.A9, Slot 29 \raggedright
\linebreak Configuration: blank. Not an access switch or distribution switch!
\linebreak Access via tsl101, like other switches.
\linebreak Name: nwt0473m66.timing.acc.gsi.de,
\linebreak IP: 192.168.21.219.
\linebreak Connections: fibre optic cable from wri2 to pexaria248t, fibre optic cable from wri3 to pexaria305t. Network cable to acc network for management.
\chapter{Software}
\section{Firmware Images}
Datamaster: build with \texttt{make ftm} in bel\_projects root folder. Current version:
\begin{verbatim}
Project     : ftm
Platform    : pexaria5 +db[12] +wrex1
FPGA model  : Arria V (5agxma3d4f27i3)
Source info : fallout-3295
Build type  : developer preview
Build date  : Wed Oct 28 16:45:33 CET 2020
Prepared by : Martin Skorsky <m.skorsky@gsi.de>
Prepared on : ACOPC042
OS version  : Linux Mint 19.3 Tricia, kernel 4.15.0-122-generic
Quartus     : Version 18.1.0 Build 625 09/12/2018 SJ Standard
                                                        Edition

  176fccd2 pmc: changed seed
  d3ab3c45 vetar2a-ee-butis: changed seed
  5419e84e microtca: changed seed
  c0d44bd6 Merge pull request #260 from
                                  GSI-CS-CO/dm-fallout-merge-v2
  50444f38 scu4: changed seed

Detecting Firmwares ...

Found 4 RAMs, 4 holding a Firmware ID


********************
* RAM @ 0x04120000 *
********************
UserLM32
Stack Status:
Project     : ftm
Version     : 7.0.1
Platform    :
Build Date  : Wed Oct 28 16:45:19 CET 2020
Prepared by : martin Martin Skorsky <m.skorsky@gsi.de>
Prepared on : ACOPC042
OS Version  : Linux Mint 19.3 Tricia  Linux 4.15.0-122-generic
                                                         x86_64
GCC Version : lm32-elf-gcc(GCC)4.5.3 (build 190527-673a32-f3d6)
IntAdrOffs  : 0x10000000
SharedOffs  : 0x500
SharedSize  : 98304
FW-ID ROM will contain:

   176fccd2 pmc: changed seed
   d3ab3c45 vetar2a-ee-butis: changed seed
   5419e84e microtca: changed seed
   c0d44bd6 Merge pull request #260 from
                                GSI-CS-CO/dm-fallout-merge-v2
   50444f38 scu4: changed seed
*****
\end{verbatim}

Pexaria for snoop: \url{https://github.com/GSI-CS-CO/bel\_projects/releases/download/fallout-v6.0.1/pexarria5.rpd}

\section{Access the Data Master}
\texttt{dm-cmd tcp/fel0069.acc.gsi.de}
\linebreak
\texttt{dm-sched tcp/fel0069.acc.gsi.de}

\section{Remote Snoop of Timing Messages}
Snoop per remote ssh: Set up of ssh without password: use public key of user@host and transfer it to root@fel0069.acc.gsi.de:~/.ssh/authorized\_keys with \texttt{ssh-copy-id}.
Snoop with Python3: the tests using python3 / pytest read the command for snooping from environment variable SNOOP.
\begin{itemize}
\item Example for local environment: saft-ctl tr0 -xv snoop 0 0 0
\item Example for remote environment on fel0069.acc.gsi.de: ssh -t root@fel0069.acc.gsi.de 'saft-ctl tr1 -xv snoop 0 0 0' \linebreak
saftd on fel0069.acc.gsi.de monitors dev/wbm1 as tr1.
\end{itemize}
The tests add an additional parameter for the number of seconds to snoop.

\section{Required versions}
\begin{enumerate}
\item Python 3.6.9
\item pytest 6.1.2
\end{enumerate}

\section{Install Image to Test on Datamaster}
This section describes the installation of \texttt{fallout-v6.1.2-rc1} on data master in test environment.
For the build, \texttt{acopc042} is used. Adopt both for a different release and build environment.

\subsection{Preparation: Build FTM Image}
\begin{verbatim}
# Preparation: quit saftd
saft-ctl tr0 quit
# switch to build folder:
cd ~/bel_projects/release
# update git repo
git pull
# switch to tagged version
git checkout fallout-v6.1.2-rc1
# build
## 1. clean up
git clean -x -f
cd ~/bel_projects/release/ip_cores/etherbone/
git clean -x -f
cd ~/bel_projects/release/ip_cores/saftlib/
git clean -x -f
cd -
git status
## 2. make and install
make
### sign kernel modules
~/scripts/sign-wishbone-modules.sh
sudo make install
### make and install etherbone
make etherbone
sudo make etherbone-install
### make and install saftlib
make saftlib
sudo make saftlib-install
sudo ldconfig
### make FTM image including FTM firmware
make ftm
# this produces:
# Project     : ftm
# Platform    : pexaria5 +db[12] +wrex1
# FPGA model  : Arria V (5agxma3d4f27i3)
## 3. Build DM Tools
cd ~/bel_projects/release/modules/ftm/ftmx86/
make
sudo make install
\end{verbatim}

\subsection{Flash on fel0069 (data master test environment)}
On \texttt{fel0069.acc.gsi.de} do (use \texttt{ssh root@fel0069.acc.gsi.de}):
\begin{verbatim}
# 1. Preparation:
Version 6.1.2 / FTM 7.0.1:
scp martin@acopc042.gsi.de:bel_projects/release/syn/gsi_pexarria5/ftm/ftm.rpd
           ftm_v6.1.2-rc1.rpd
wget https://github.com/GSI-CS-CO/bel_projects/releases/download
            /fallout-v6.1.2-rc1/falloutv6_1_2-rc1-pexarria5.rpd
Version 6.2.0 / FTM 8.0.0:
scp martin@acopc042.gsi.de:Documents/fel0069/ftm_722413ec_2021_11_02.rpd
           ftm_722413ec_2021_11_02.rpd
pexarria5: same as version 6.1.2.
## stop fesa: how to?
## stop saftd - error message if saftd is not running
saft-ctl tr1 quit
## stop datamaster
dm-cmd dev/wbm0 halt
## halt all lm32 CPUs on both pexarrias
eb-reset dev/wbm0 cpuhalt 0xff
eb-reset dev/wbm1 cpuhalt 0xff
# 2. Flash - don't use eb-flash-secure.
# Included preparation not ok.
# Parameters -s and -w not supported
eb-flash -s 0x10000 -w 3 dev/wbm0 'ftm rpd file'
eb-flash -s 0x10000 -w 3 dev/wbm1 'pexarria5 rpd file'
# 3. Reboot
eb-reset dev/wbm0 fpgareset; eb-reset dev/wbm1 fpgareset; reboot
\end{verbatim}
Now, if result of flash looks OK, update PXE boot (on ASL cluster).
\begin{verbatim}
ssh mskorsky@asl755.acc
cd /common/export/nfsinit/fel0069/
# adopt gateware timing rte
rm 20_timing-rte
ln -s ../global/timing-rte-tg-fallout-v6.1.2-rc1 20_timing-rte
# adopt DM tools, includes dm-cmd, dm-sched, libcarpedm
rm 40_generator_dev
ln -s ../global/dm-.... 40_generator_dev
str-D # end ssh session
\end{verbatim}
\begin{verbatim}
ssh root@fel0069.acc
reboot
\end{verbatim}
Finally: check that datamaster is up and run tests (do this on acopc042).
\begin{verbatim}
dm-cmd tcp/fel0069.acc
\end{verbatim}
Run tests on tsl021.acc.gsi.de: start Jenkins Jobs \url{https://builder.acc.gsi.de/jenkins/job/timing/view/tests/job/test_module_ftm_datamaster_dm-fallout-tests/}

\end{document}
